Mining Interaction Log Data in a Creative Arts Practice using Macro Gestures and Transition Matrices

### 2015-09-20 flux-entropy-paper/flux-entropy-paper.tex
chucked out nonperformative data from plots

%   %In this note we describe a data archive for an iPad
% musical-ensemble performance practice which has included concerts,
% demonstrations and formal HCI studies of groups of musicians over more
% than two years. Our corpus contains data for 150 performances,
% rehearsals, and demonstrations, involving some 100 different musicians
% and several different iPad apps. At the lowest level, data for each
% musician in a particular performance was archived as Apple iOS
% callback methods for touching and swiping. Subsequent data analysis
% has been undertaken by first classifying these data into higher level
% gesture types and then describing the transitions between these
% gestures for each artist as a function of time. We find that a
% transition-matrix measure, which we call ``flux'', can be used to
% distinguish different performance types and contexts, as well as
% different phases (beginning, middle and end) of all of the
% performances in our corpus.

@@

\category{H.5.5.}{Information Interfaces and Presentation
  (e.g. HCI)}{Sound and Music Computing}{Systems}

@@

A useful way of modelling user interfaces is to consider their configurations as state vectors and to deploy the methods of matrix algebra to describe transitions between these states during an interaction sequence\cite{Thimbleby:2001kq, Thimbleby:2004fj}. Data obtained from usability analysis of quite complex systems (such as a resource planning application for anti-terrorism planning [ref: Kannampallil, Thomas, and Steven Haynes. "Using Event Based Markov Model Simulation to Analyze Interactive Human Behavior." AMCIS 2007 Proceedings (2007): 142. ]) have been undertaken by coding sequences of fine-detail human actions as states and then to model data from usability walkthroughs as sequences of state transitions. The transition probably matrix for transitions between states can then be derived from such user data and conclusions drawn about the nature of interaction. In the creative arts, editing commands at a computer interface for the ``live coding of computer music'' have also been coded coded as states and analysed using transition matrices, in both the textual and musical dimensions, in order to arrive at conclusions regarding artistic style\cite{Swift:2014tya}.

In our corpus of iPad musical performances, rehearsals and studies, the macro gestures identified in [ref: charlesHCI] provide a basis to identify and code macro gestures. Each performer in each muscial piece can then be thought of as following a ``score'' of state transitions and an overall transition matrix for the entire ensemble can be derived by averaging the individual transition matrices.

@@

In the rest of this paper we will use these matrix measures to answer
two questions. First, do the transition matrix measures allow us to
distinguish between different performance contexts, types and
instrumental setups? Secondly, do these performance measure allow us
further insight into the corpus than the metadata can provide? TODO I
think this second question is too broad?

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/flux-entropy-distribution}
%   \caption{Distribution of flux and entropy by performance type and
%     context. Flux and entropy tended to be linearly related in
%     improvisations, particularly in demonstrations and studies, while
%     compositions had a more scattered relationship.
%     \label{fig:flux-entropy-distribution}}
% \end{figure}

@@

\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}

>>>>

  In this note we describe the analysis of a data archive for an iPad
  musical-ensemble performance practice where transition matrix
  measures are able to distinguish between sessions of different
  contexts, and the internal structure of these sessions. The archive
  includes 95 touch data logs of ensemble touch-screen interactions in
  concerts, rehearslas and formal HCI studies of groups of musicians
  over more than two years.

  % At the lowest level, data for each musician in a particular
  % performance was archived as Apple iOS callback methods for
  % touching and swiping.

  The data analysis has been undertaken by first classifying
  these data into higher level gesture types and then describing the
  transitions between these gestures for each artist as a function of
  time. We find that a transition-matrix measure, which we call
  ``flux'', can be used to distinguish different performance types and
  contexts, as well as different phases (beginning, middle and end) of
  all of the performances in our corpus.

@@

\category{H.5.5.}{Information Interfaces and Presentation
  (e.g. HCI)}{Sound and Music Computing}{ Systems}

@@

A useful way of modelling user interfaces is to consider their
configurations as state vectors and to deploy the methods of matrix
algebra to describe transitions between these states during an
interaction sequence\cite{Thimbleby:2001kq, Thimbleby:2004fj}. Data
obtained from usability analysis of quite complex systems (such as 
resource planning software for security applications~\cite{Kannampallil:2007fp})

have been undertaken by coding sequences of fine-detail human actions
as states and then to model data from usability walkthroughs as
sequences of state transitions. The transition probably matrix for
transitions between states can then be derived from such user data and
conclusions drawn about the nature of interaction. In the creative
arts, editing commands at a computer interface for the ``live coding
of computer music'' have also been coded coded as states and analysed
using transition matrices, in both the textual and musical dimensions,
in order to arrive at conclusions regarding artistic
style\cite{Swift:2014tya}.

In our corpus of iPad musical performances, rehearsals and studies,
the macro gestures identified by Martin et al~\cite{Martin:2014cr}
provide a basis to identify and code macro gestures. Each performer in
each muscial piece can then be thought of as following a ``score'' of
state transitions and an overall transition matrix for the entire
ensemble can be derived by averaging the individual transition
matrices.

@@

In the rest of this paper we will use these matrix measures to answer
two questions. First, do the transition matrix measures allow us to
distinguish between different performance contexts, types and
instrumental setups? Secondly, do these performance measure allow us
further insight into the corpus than the metadata can provide?
% TODO I think this second question is too broad?

@@

@@

%\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}
